---
title: "Brain age supplemental tables"
---

Format and output supplemental tables to paste in


```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(boot)
library(cowplot)
library(glmnet)
library(kableExtra)
library(broom)

```

Set the filepaths for input and output
ENSURE THIS IS A LOCATION SUITABLE FOR STORAGE OF UKB DATA
```{r}

input_path <- # data set storage location

output_path <- # script outputs storage location

```

Put the date IDPs processed
```{r}

run_date <- "2024-11-13" 

```

Neurological exclusions
```{r}

load(paste0(output_path, "neuro_exclusions_", run_date, ".RData"))

```


```{r}

neuro_disease_table

```


```{r}

write.csv(neuro_disease_table, file = paste0("neuro_exclusions_", Sys.Date(), ".csv"))

```




Missing report

```{r}


load(paste0(output_path, "missing_report_p5_", run_date, ".RData"))


```

```{r}

missing_report <- missing_report %>%
  pivot_wider(id_cols = data_field, names_from = step, values_from = n_missing) %>%
  select(-c(Import, 'Manual BP')) %>%
  mutate(data_field = factor(data_field, levels = c("sex", "age", "week_drinks", "pack_years",
                                                    "income", "education_years", "BMI", "diabetes", "systolic_BP",
                                                    "diastolic_BP", "health_rating", "longstanding",
                                                    "head_size", "site", "date", "rfMRI_motion", "genetic_PC"))) %>%
  arrange(data_field) %>%
  filter(! is.na(data_field)) %>%
  rbind(data.frame("data_field" = c("genetic_PC"),
                   "Before" = c(0),
                   "After" = c(0))) %>%
  mutate(After = ifelse(is.na(After), 0, After))

missing_report

```

Save as .csv for pasting
```{r}

write.csv(missing_report, file = paste0("formatted_missing_report_", Sys.Date(), ".csv"))

```







load(paste0(output_path, "neuro_exclusions_", run_date, ".RData"))
write.csv(missing_report, file = paste0("formatted_missing_report_", Sys.Date(), ".csv"))





