---
title: "Category regressions with GMV only"
---

Run alcohol only and alcohol + brainPAD but with the category regression instead of continuous

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(kableExtra)

```

```{r}

model_version <- "main"

```

Set the filepaths for input and output
ENSURE THIS IS A LOCATION SUITABLE FOR STORAGE OF UKB DATA
```{r}

input_path <- # data set storage location

output_path <- # script outputs storage location

```


Put the date IDPs processed
```{r}

run_date <- "2024-11-13" 

```


```{r}

load(paste0(output_path, "IDP_names_table_", run_date, ".RData"))

```


Load genetic PCs
```{r}

load(paste0(output_path, "genetic_PC_", run_date, ".RData"))

```

Load the processed imaging covariates
From step 2 of the pipeline
```{r}

load(paste0(output_path, "/processed_imaging_covariates_", "2024-07-29", ".RData"))

```

Load the covariates that have been processed with backfilling imaging data with baseline data and then imputation
From step 3 of the pipeline
```{r}

load(paste0(output_path, "/imputed_covariates_", run_date, ".RData"))

```

```{r}

load(paste0(output_path, model_version, "_brainPAD_", run_date, ".RData"))

```


Combine the tables to make a regression table
Create a column for scaled brainPAD
```{r}

regression_table <- imputation_results %>%
  select(-c(sex, age, date)) %>%
  filter(n_eid %in% experimental_brainPAD$n_eid) %>%
  left_join(genetic_PC, by = "n_eid") %>%
  left_join(imaging_covariates, by = "n_eid") %>%
  left_join(select(experimental_brainPAD, -age), by = "n_eid") %>%
  mutate(scaled_brainPAD = scale(brainPAD, center = TRUE, scale = TRUE))
  
rm(imputation_results)
rm(imaging_covariates)
rm(genetic_PC)
gc()

```

Create the alcohol category column
Use the week drinks variable in the covariates to create bins
If changing this make sure to also change in demographics or the numbers will not line up!
```{r}

regression_table <- regression_table %>%
  mutate(drink_group = case_when(
    alcohol_status == "Never" ~ 6,
    alcohol_status == "Former" ~ 6,
    week_drinks <= 1 ~ 0, # Current alcohol consumption with less than 1 drink per week are the reference group
    week_drinks <= 7 ~ 1,
    week_drinks <= 14 ~ 2,
    week_drinks <= 21 ~ 3,
    week_drinks <= 28 ~ 4,
    week_drinks > 28 ~ 5))


```


Note that the < 1 group includes some participants who indicated they do drink (current) in data field 1558 but who end up with calculated week drinks of 0

Factor drink groups, and specify that 0 (lightest drinking group) is the reference level
```{r}

regression_table$drink_group <- factor(regression_table$drink_group)
#regression_table$drink_group <- relevel(regression_table$drink_group, ref = 0)

class(regression_table$drink_group)
levels(regression_table$drink_group)

```

Check the N in each drinking category
```{r}

group_n <- regression_table %>%
  group_by(drink_group) %>%
  tally()

group_n

```



Scale brainPAD
```{r}

regression_table <- regression_table %>%
  mutate(scaled_brainPAD = scale(brainPAD, center = TRUE, scale = TRUE))

```


Set up the list of covariates (updated 7/25/24)
```{r}

covariates <-c("sex", "age", "scaled_pack_years", "income", "education_years",
               "BMI", "diabetes", "systolic_BP", "diastolic_BP", "health_rating", "longstanding",
               "head_size", "site", "date", "rfMRI_motion",
               "PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6", "PC_7", "PC_8", "PC_9", "PC_10")


predictor <- "drink_group"

```


```{r}

regression <- glm(formula(paste("regression_table$brainPAD ~",
                                             paste(c(predictor, covariates), collapse = "+"))),
                                              family = "gaussian", data = regression_table)

```

Clean up the regression results
```{r}

result_table <- tidy(regression)
result_table

```

Make the results into a nice table showing the relationship

```{r}

brain_PAD_regression <- result_table %>%
  filter(grepl("drink_group", term)) %>%
  # Add a row to account for the reference group (infrequent, < 1)
  rbind(data.frame("term" = c("0"), "estimate" = c(0), std.error = c(NA), statistic = c(NA), "p.value" = c(NA))) %>%
  mutate(drink_group = gsub("drink_group", "", term)) %>%
  # Factor and order
  mutate(drink_group = factor(drink_group, levels = c(6, 0, 1, 2, 3, 4, 5))) %>%
  arrange(drink_group) %>%
  # Add readable group names
  mutate(week_drinks = case_when(
    drink_group == 6 ~ "0 (never)",
    drink_group == 0 ~ "<=1*",
    drink_group == 1 ~ ">1-7",
    drink_group == 2 ~ ">7-14",
    drink_group == 3 ~ ">14-21",
    drink_group == 4 ~ ">21-28",
    drink_group == 5 ~ ">28")) %>%
  left_join(group_n, by = "drink_group") %>%
  # Calculate confidence intervals
  mutate(interval = 1.86 * std.error) %>%
  mutate(lower = estimate - interval) %>%
  mutate(upper = estimate + interval) %>%
  select(-c(term, statistic, std.error, drink_group, interval)) %>%
  relocate(estimate, .after = n) %>%
  relocate(p.value, .after = upper)

brain_PAD_regression

```

```{r}

theme_replace(
  # Set the axis text size
  axis.text.x = element_text(size = 13), axis.text.y = element_text(size = 13),
  # Background white with no grid lines
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = "white", color = "dark grey"))

```


Make it into a figure
```{r}

brain_PAD_regression %>%
  mutate(week_drinks = factor(week_drinks, levels = c("0 (never)", "<=1*", ">1-7", ">7-14",
                                                      ">14-21", ">21-28", ">28"))) %>%
  ggplot(aes(x = week_drinks, y = estimate)) +
  geom_point(size = 1.8, color = "blue") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, size = 0.8) +
  ylab("BrainPAD (years)") +
  xlab("Drinking category (drinks per week)") +
  geom_hline(yintercept = 0, color = "darkgrey", size = 0.8, alpha = 0.5) +
  ylim(c(-1, 3))

```



