---
title: "Category regressions with GMV only"
---

Run alcohol only and alcohol + brainPAD but with the category regression instead of continuous

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(kableExtra)

```

```{r}

model_version <- "main"

```

Set the filepaths for input and output
ENSURE THIS IS A LOCATION SUITABLE FOR STORAGE OF UKB DATA
```{r}

input_path <- # data set storage location

output_path <- # script outputs storage location

```

Put the date IDPs processed
```{r}

run_date <- "2024-11-13" 

```


```{r}

load(paste0(output_path, "IDP_names_table_", run_date, ".RData"))

```


Load genetic PCs
```{r}

load(paste0(output_path, "genetic_PC_", run_date, ".RData"))

```

Load the processed imaging covariates
From step 2 of the pipeline
```{r}

load(paste0(output_path, "/processed_imaging_covariates_", "2024-07-29", ".RData"))

```

Load the covariates that have been processed with backfilling imaging data with baseline data and then imputation
From step 3 of the pipeline
```{r}

load(paste0(output_path, "/imputed_covariates_", run_date, ".RData"))

```

```{r}

load(paste0(output_path, model_version, "_brainPAD_", run_date, ".RData"))

```


Combine the tables to make a regression table
Create a column for scaled brainPAD
```{r}

regression_table <- imputation_results %>%
  select(-c(sex, age, date)) %>%
  filter(n_eid %in% experimental_brainPAD$n_eid) %>%
  left_join(genetic_PC, by = "n_eid") %>%
  left_join(imaging_covariates, by = "n_eid") %>%
  left_join(select(experimental_brainPAD, -age), by = "n_eid") %>%
  mutate(scaled_brainPAD = scale(brainPAD, center = TRUE, scale = TRUE))
  
rm(imputation_results)
rm(imaging_covariates)
rm(genetic_PC)
gc()

```

Create the alcohol category column
Use the week drinks variable in the covariates to create bins
If changing this make sure to also change in demographics or the numbers will not line up!
```{r}

regression_table <- regression_table %>%
  mutate(drink_group = case_when(
    alcohol_status == "Never" ~ 6,
    alcohol_status == "Former" ~ 6,
    week_drinks <= 1 ~ 0, # Current alcohol consumption with less than 1 drink per week are the reference group
    week_drinks <= 7 ~ 1,
    week_drinks <= 14 ~ 2,
    week_drinks <= 21 ~ 3,
    week_drinks <= 28 ~ 4,
    week_drinks > 28 ~ 5))


```


Note that the < 1 group includes some participants who indicated they do drink (current) in data field 1558 but who end up with calculated week drinks of 0

Factor drink groups, and specify that 0 (lightest drinking group) is the reference level
```{r}

regression_table$drink_group <- factor(regression_table$drink_group)
#regression_table$drink_group <- relevel(regression_table$drink_group, ref = 0)

class(regression_table$drink_group)
levels(regression_table$drink_group)

```

Check the N in each drinking category
```{r}

group_n <- regression_table %>%
  group_by(drink_group) %>%
  tally()

group_n

```


Define a function for regression (will run on all the different sets of IDPs)

```{r}


do_IDP_regression <- function(IDP_table, vars_list, regression_table) {
  # Function to perform regression using covariates in the regression table on all the IDPs in the supplied IDP table
  # Will return a table with IDP, alcohol estimate and P value
  
  # Join the IDPs onto the regression table
  # Changed this to inner join, R has been sneakily dropping my n the WHOLE time
  regression_table <- inner_join(IDP_table, regression_table, by = "n_eid", all = FALSE)
  regression_table <- na.omit(regression_table)
  # Print the N so user is alerted if this happens
  print(paste0("IDP N=", nrow(IDP_table), ", regression N=", nrow(regression_table)))
  
  IDP_names <- colnames(IDP_table)
  IDP_names <- IDP_names[IDP_names != "n_eid"]
  
  regression <- lapply(IDP_names, function(x) glm(formula(paste("unlist(regression_table[,x]) ~",
                                             paste(vars_list, collapse = "+"))),
                                              family = "gaussian", data = regression_table))
  regression <- lapply(regression, tidy)
  
  regression_results <- data.frame("alcohol_beta" = c(), "alcohol_p" = c())

  for (n in 1:length(regression)) {
    
    IDP <- IDP_names[n]
    summary_tibble <- regression[[n]]
    
    summary_tibble <- summary_tibble %>%
      mutate(term = gsub("\\(", "", term)) %>%
      mutate(term = gsub("\\)", "", term)) %>%
      mutate(term = tolower(term)) %>%
      select(term, estimate, std.error, p.value) %>%
      pivot_longer(cols = c(estimate, std.error, p.value), names_to = "stat", values_to = "value")
    
    summary_tibble <- summary_tibble %>%
      mutate(col_name = case_when(
      (stat == "estimate") ~ paste0(term, "_beta"),
      (stat == "std.error") ~ paste0(term, "_se"),
      (stat == "p.value") ~ paste0(term, "_p"))) %>%
      select(col_name, value) %>%
      pivot_wider(names_from = col_name, values_from = value)
    
    new_row <- data.frame("IDP" = IDP)
    
    new_row <- cbind(new_row, summary_tibble)
    
    
    regression_results <- rbind(regression_results, new_row)
    
  }
  

  return(regression_results)
}


```


Scale brainPAD
```{r}

regression_table <- regression_table %>%
  mutate(scaled_brainPAD = scale(brainPAD, center = TRUE, scale = TRUE))

```


Set up the list of covariates (updated 7/25/24)
```{r}

covariates <-c("sex", "age", "scaled_pack_years", "income", "education_years",
               "BMI", "diabetes", "systolic_BP", "diastolic_BP", "health_rating", "longstanding",
               "head_size", "site", "date", "rfMRI_motion",
               "PC_1", "PC_2", "PC_3", "PC_4", "PC_5", "PC_6", "PC_7", "PC_8", "PC_9", "PC_10")

```


Load IDPs and perform regression
```{r}

total_volume_IDPs <- c("X26514.2.0", "X26518.2.0", "X26527.2.0", "total_wm")


load(paste0(output_path, "ukbio_IDP_", run_date, ".RData"))

ukbio_IDP <- ukbio_IDP %>%
  select(c(n_eid, all_of(total_volume_IDPs))) %>%
  # NORMALIZE
  mutate_at(vars(-n_eid), ~scale(.x, center = TRUE, scale = TRUE))

```


```{r}

predictors <- c("drink_group")

```


```{r}

alcohol_model <- do_IDP_regression(ukbio_IDP, c(predictors, covariates), regression_table)

```


=================================================================
Alcohol and brainPAD model

```{r}

predictors <- c("drink_group", "scaled_brainPAD")

```


```{r}

alcohol_brainPAD_model <- do_IDP_regression(ukbio_IDP, c(predictors, covariates), regression_table)

```

Make a table with just GMV from Freesurfer ASEG
X26518.2.0 ~ total grey matter volume

```{r}

GMV_alcohol <- alcohol_model %>%
  filter(IDP == "X26518.2.0") %>%
  select(-IDP) %>%
  pivot_longer(everything(), names_to = "covariate", values_to = "value") %>%
  dplyr::rename("alcohol" = value)

GMV_alcohol_brainPAD <- alcohol_brainPAD_model %>%
  filter(IDP == "X26518.2.0") %>%
  select(-IDP) %>%
  pivot_longer(everything(), names_to = "covariate", values_to = "value") %>%
  dplyr::rename("with_brainPAD" = value)

GMV_table <- GMV_alcohol %>%
  right_join(GMV_alcohol_brainPAD, by = "covariate")

```

Save the regression results - NO IDP info
```{r}

#save(alcohol_model, file = paste0(output_path, model_version, "_alcohol_model_", Sys.Date(), ".RData"))

#save(alcohol_brainPAD_model, file = paste0(output_path, model_version, "_alcohol_brainPAD_model_", Sys.Date(), ".RData"))

```

Make the table for GMv before and after controlling for brainPAD
```{r}

GMV_table <- GMV_table %>%
  filter(grepl("drink_group", covariate)) %>%
  mutate(drink_group = gsub("drink_group", "", covariate)) %>%
  separate(col = drink_group, into = c("drink_group", "measure"), sep = "_") %>%
  select(-covariate) %>%
  filter(measure != "se") %>%
  pivot_longer(cols = -c(drink_group, measure), names_to = "model", values_to = "value") %>%
  mutate(col_name = paste0(model, "_", measure)) %>%
  arrange(model) %>%
  select(-c(measure, model)) %>%
  pivot_wider(id_cols = drink_group, names_from = col_name, values_from = value) %>%
  # Add a row to account for the reference group (infrequent, < 1)
  rbind(data.frame("drink_group" = c("0"), "alcohol_beta" = c(0), "alcohol_p" = c(NA),
                   "with_brainPAD_beta" = c(0), "with_brainPAD_p" = c(NA))) %>%
  mutate(drink_group = factor(drink_group, levels = c(6, 0, 1, 2, 3, 4, 5))) %>%
  arrange(drink_group) %>%
  # Add readable group names
  mutate(week_drinks = case_when(
    drink_group == 6 ~ "0 (never)",
    drink_group == 0 ~ "<=1*",
    drink_group == 1 ~ ">1-7",
    drink_group == 2 ~ ">7-14",
    drink_group == 3 ~ ">14-21",
    drink_group == 4 ~ ">21-28",
    drink_group == 5 ~ ">28")) %>%
  select(-drink_group) %>%
  relocate(week_drinks, .before = alcohol_beta) %>%
  mutate_at(vars(alcohol_beta, with_brainPAD_beta), ~formatC(.x, digits = 2, format = "f")) %>%
  mutate_at(vars(alcohol_p, with_brainPAD_p), ~ifelse(.x > 0.009, formatC(.x, digits = 2, format = "f"),
                                                           formatC(.x, digits = 2, format = "e")))

GMV_table

```

Save this as a nice .csv for the paper
```{r}

write.csv(GMV_table, file = paste0("GMV_category_table_", Sys.Date(), ".csv"))

```














